Left-linear rules in adhesive categories

### 2024-02-13 icalp/fusione.tex
minima

\begin{abstract}
%  When two sequences of steps executed by a computational device can
%  be considered equivalent?  This is a relevant question if one
%  considers concurrent and distributed systems, since the independence
%  of two steps means that they could be executed in any order, and
%  potentially in parallel.
  When can two sequential steps performed by a computing device be considered (causally) independent? 
  This is a relevant question for concurrent and distributed systems, since independence
  means that they could be executed in any order, and potentially in parallel.
  % 
  We investigate the issue in the context of adhesive categories, a
  very general instance of the DPO approach to rewriting. More
  precisely, we focus on left-linear rules, which not only consume,
  preserve and generate entities, as linear rules do, but may
  ``merge'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box, but it requires the
  restriction to well-behaved classes of left-linear rules.
  %
  Moreover, a standard characterisation of independence that -- in the
  linear case -- allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds.
  %and this requires a substantial revision of the theory.
  %
  The paper performs an in-depth study of the notion of independence
  for left-linear rules: it introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent computations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}

@@

So far, most of the theoretical results focussed on what are called \emph{linear rules}: 
intuitively, a step is required only to consume, preserve and generate entities. 
However, in some situations it is also necessary to ``merge'' parts of the state. 
This happens already in the context of string diagram mentioned above, but also 
in graphical implementations of nominal calculi where, as a result of name passing, 
the received name is identified with a local one at the
receiver~\cite{CVY:ESSPE,Gad07} or in the visual modelling of bonding in
biological/chemical processes~\cite{PUY:MBPE}. A similar mechanisms is
at the core of e-graphs (equality graphs), recently used for
rewrite-driven compiler optimisations~\cite{WNW:egg}: rather than
explicitly merging nodes corresponding to equal expression, the idea
here -- conceptually and, as we will see, also mathematically similar --
is to maintain an equivalence %\todo{P: more precisely a congruence}
over the nodes.  
%\todo{P: la sequenza di lavori su questo \`e ampia,
%vedere se c'\`e qualcosa d'altro}

Technically, to acquire the expressiveness needed for capturing
fusions, i.e. the merge of some elements in the state, requires to
move from linear to left-linear rewriting rules. While left-linear
rules have been considered in various instances (mainly in categories
of graphs) and some results concerning the corresponding rewriting
theory have been put forward, to the best of our knowledge the
phenomena arising when devising a theory of indepedence and
concurrency for rewriting with left-linear rules in the general
setting of adhesive and quasi-adhesive categories
%(which capture most ``applications'') 
have not been systematically
explored~\cite{Ehrig1976,EHP:BRfTToHLRS,baldan2011adhesivity} \todo{P: see also \cite{HMP:DPORADD,BHK:CTNL} ... verificare e aggiungere}

Here we perform this exploration working at the level of $\mathcal{M}$-adhesive categories, 
%i.e.such that adhesivity holds only for a class $\mathcal{M}$ of arrows, 
which encompass a large family of DPO-based formalisms adopted by pratictioners~\cite{xxx}.

@@

There is, however, a further problem.  For linear
rewriting systems, switch equivalence can be characterised in terms of the
existence of what is called a consistent permutation~\cite{xxx}.
%(roughly, an isomorphism between their colimits). 
This result, which
is folklore for various categories of graphs, is explicitly proved
here for $\mathcal{M}$-adhesive categories.
%
The characterisation of switch equivalence based on consistent
permutations is the key to derive some
properties for linear rewriting systems, namely
\begin{enumerate}
\item \emph{consistency of switching}: when transforming
  sequences of rewriting steps into equivalent ones by switching independent steps, the
  result not depends on the order in which switchings are applied, but
  only on the resulting permutation;
\item \emph{globality of independence}: if two independent steps are
  moved forward or backward due to the application of switchings to a
  derivation, they remain independent.
\end{enumerate}
Properties (1) and (2) allow one to deduce a canonical form for equivalence proofs 
between sequences of rewriting steps seen as concurrent computations.

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails: the
existence of a consistent permutation is only necessary but not
sufficient, hence it cannot be used for proving (1) and (2).
%
This leads to the question asking whether the properties above hold.
We show that actually much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weak form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality. Finally, we prove that a canonical form
for switching sequences can also be recovered.

@@

\begin{definition}[well switching rewriting systems]
A left-linear DPO rewriting system $(\X, \R)$ is \emph{well switching} if it is strong enforcing and, for every derivation $\der{D}:=\{\dder{D}_{i}\}_{i=0}^1$, there is at most one independence paire between $\dder{D}_0$ and $\dder{D}_1$.
\end{definition}

@@

\begin{restatable}[Globality of independence]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a strong enforcing DPO rewriting system.
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is an independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item
    \label{lem:indep-global-left:1}
    $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
    switchable.

  \item
    \label{lem:indep-global-left:2}
    If $S_{i_0, i_1}(\dder{D}_0)\seqind \dder{D}_2$ with an independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)\seqind S_{\alpha_0, \alpha_1}(\dder{D}_2)$.
  \end{enumerate}
\end{restatable}

@@

	\begin{figure}[h]
		\centering
	$\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){i_0}|(.7)\hole
		\ar[dr]^{h_0}&& L_1\ar@/^.8cm/[ddll]^(.2){i_1}
		\ar[dl]_{m_1}\\&K_0\ar[dr]^{k_0}\ar[dl]_{l_0}
		\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){v_0}\ar[ur]^{r_0}&& G_1 &&
		K_1
		\ar@/^.6cm/[ddll]|(.36)\hole_(.65){v_1}\ar[dl]_{k_1}\ar[lu]_{l_1}
		\ar[dr]^{r_1}\\L_0
		\ar@/_.6cm/[ddrr]_(.2){j_0}|(.31)\hole|(.81)\hole
		\ar[dr]^(.4){m_0}|(.61)\hole && D_0
		\ar[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&D_1
		\ar[dr]|(.4)\hole^(.65){g_1}
		\ar[ul]|(.5)\hole_(.65){f_1}&&R_1\ar@/^.6cm/[ddll]^(.2){j_1}|(.31)\hole|(.81)\hole\ar[dl]_(.4){h_1}|(.6)\hole\\&G_0
		&&P_1\ar[dr]_{q_1}
		\ar[dl]^{q_0}\ar[ur]^(.4){p_1}\ar[ul]_(.4){p_0}&&G_2\\L_1
		\ar@/^.6cm/[uurr]^(.2){i_1} \ar[ur]_(.35){f_0\circ
			i_1}|(.61)\hole&&Q_0
		\ar[ul]|(.4)\hole_(.65){s_0}\ar[dr]|(.5)\hole^(.65){t_0}
		&&Q_1\ar[ur]|(.4)\hole^(.65){t_1} \ar[dl]|(.5)\hole_(.65){s_1}
		&& R_0 \ar[ul]^(.35){g_1\circ
			i_0}|(.61)\hole\ar@/_.6cm/[uull]_(.2){i_0}\\&K_1
		\ar[ur]_{q_0\circ v_1} \ar[dr]_{r_1}
		\ar[ul]^{l_1}\ar@/^.6cm/[uurr]_(.65){v_1}&&H'_1&& K_0
		\ar[ur]_{r_0} \ar[ul]^{q_1\circ v_0} \ar[dl]^{l_0}
		\ar@/_.6cm/[uull]^(.65){v_0}\\&& R_1
		\ar[ur]_{\hspace{-6pt}s_1\circ
			j_1}\ar@/^.8cm/[uurr]^(.2){j_1}&& L_0 \ar[ul]^{t_0\circ
			j_0\hspace{-6pt}} \ar@/_.8cm/[uull]_(.2){j_0} |(.69)\hole
	}	$
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switching of $\dder{D}_0$ and $\dder{D}_1$}
		\label{fi:first}
	\end{figure}

	\begin{figure}[h]
		\centering 
		$\xymatrix@C=15pt@R=15pt{&&R_1 \ar@/_.8cm/[ddrr]_(.2){a_0}|(.7)\hole
			\ar[dr]^{h_1}&& L_2\ar@/^.8cm/[ddll]^(.2){a_1}
			\ar[dl]_{m_2}\\&K_1\ar[dr]^{k_1}\ar[dl]_{l_1}
			\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){u_1}\ar[ur]^{r_1}&& G_2 &&
			K_2
			\ar@/^.6cm/[ddll]|(.36)\hole_(.65){u_2}\ar[dl]_{k_2}\ar[lu]_{l_2}
			\ar[dr]^{r_2}\\L_1
			\ar@/_.6cm/[ddrr]_(.2){b_0}|(.31)\hole|(.81)\hole
			\ar[dr]^(.4){m_1}|(.61)\hole && D_1
			\ar[dl]|(.4)\hole_(.65){f_1}\ar[ur]|(.5)\hole^(.7){g_1}&&D_2
			\ar[dr]|(.4)\hole^(.65){g_2}
			\ar[ul]|(.5)\hole_(.65){f_2}&&R_2\ar@/^.6cm/[ddll]^(.2){b_1}|(.31)\hole|(.81)\hole\ar[dl]_(.4){h_2}|(.61)\hole\\&G_1
			&&P_2\ar[dr]_{d_1}
			\ar[dl]^{d_0}\ar[ur]^(.4){c_1}\ar[ul]_(.4){c_0}&&G_3\\L_2
			\ar@/^.6cm/[uurr]^(.2){a_1} \ar[ur]_(.35){f_1\circ
				a_1}|(.61)\hole&&Q_2
			\ar[ul]|(.4)\hole_(.65){x_1}\ar[dr]|(.5)\hole^(.65){y_1}
			&&Q_3\ar[ur]|(.4)\hole^(.65){y_2} \ar[dl]|(.5)\hole_(.65){x_2}
			&& R_1 \ar[ul]^(.3){g_2\circ
				a_0}|(.61)\hole\ar@/_.6cm/[uull]_(.2){a_0}\\&K_2
			\ar[ur]_{d_0\circ u_2} \ar[dr]_{r_2}
			\ar[ul]^{l_2}\ar@/^.6cm/[uurr]_(.65){u_2}&&G'_2&& K_1
			\ar[ur]_{r_1} \ar[ul]^{d_1\circ u_1} \ar[dl]^{l_1}
			\ar@/_.6cm/[uull]^(.65){u_1}\\&& R_2
			\ar[ur]_{\hspace{-6pt}x_2\circ
				b_1}\ar@/^.8cm/[uurr]^(.2){b_1}&& L_1 \ar[ul]^{y_1\circ
				b_0\hspace{-7pt}} \ar@/_.8cm/[uull]_(.2){b_0} |(.69)\hole
		}	$		
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switching of $\dder{D}_1$ and $\dder{D}_2$}
		\label{fi:second}
	\end{figure}

	\begin{figure}
		\centering
		$\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){e_0}|(.7)\hole
			\ar[dr]^{\hspace{-5pt}h_0}&& L_2\ar@/^.8cm/[ddll]^(.2){e_1}
			\ar[dl]_{f_1\circ a_1\hspace{-5pt}}\\&K_0\ar[dr]^{k_0}\ar[dl]_{l_0}
			\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){w_0}\ar[ur]^{r_0}&& G_1 &&
			K_2 \ar@/^.6cm/[ddll]|(.36)\hole_(.65){w_2}\ar[dl]_{d_0\circ
				u_2\hspace{-5pt}}\ar[lu]_{l_2} \ar[dr]^{r_2}\\L_0
			\ar@/_.6cm/[ddrr]_(.2){o_0}|(.31)\hole|(.81)\hole
			\ar[dr]^(.4){m_0}|(.61)\hole && D_0
			\ar[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&Q_2
			\ar[dr]|(.4)\hole^(.65){y_1}
			\ar[ul]|(.5)\hole_(.65){x_1}&&R_2\ar@/^.6cm/[ddll]^(.2){o_1}|(.31)\hole|(.81)\hole\ar[dl]_(.3){x_2\circ
				b_1}|(.58)\hole \\&G_0 &&P_3\ar[dr]_{n_1}
			\ar[dl]^{n_0}\ar[ur]^(.4){u_1}\ar[ul]_(.4){u_0}&&G'_2\\L_2
			\ar@/^.6cm/[uurr]^(.2){e_1} \ar[ur]_(.35){f_0\circ
				e_1}|(.61)\hole&&Q_4
			\ar[ul]|(.4)\hole_(.65){z_0}\ar[dr]|(.5)\hole^(.65){z'_0}
			&&Q_5\ar[ur]|(.4)\hole^(.65){z'_1}
			\ar[dl]|(.5)\hole_(.65){z_1} && R_0 \ar[ul]^(.3){y_1\circ
				e_0}|(.6)\hole\ar@/_.6cm/[uull]_(.2){e_0}\\&K_2
			\ar[ur]_{\hspace{-2pt}n_0\circ w_2} \ar[dr]_{r_2}
			\ar[ul]^{l_2}\ar@/^.6cm/[uurr]_(.65){w_2}&&G'_1&& K_0
			\ar[ur]_{r_0} \ar[ul]^{n_1\circ w_0} \ar[dl]^{l_0}
			\ar@/_.6cm/[uull]^(.65){w_0}\\&& R_2
			\ar[ur]_{\hspace{-4pt}z_1\circ
				o_1}\ar@/^.8cm/[uurr]^(.2){o_1}&& L_0 \ar[ul]^{z'_0\circ
				o_0\hspace{-7pt}} \ar@/_.8cm/[uull]_(.2){o_0} |(.69)\hole
		}	$
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switching of $\dder{D}_0$ and $S_{a_0,a_1}(\dder{D}_2)$}
		\label{fi:third}
	\end{figure}

	\begin{enumerate}
		\item We have to construct the two dotted arrows in the diagram
		below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0
			\ar[d]_{n_1\circ w_0}\ar[ll]_{l_0} \ar[r]^{r_0} & R_0
			\ar@{.>}@/^.35cm/[drrr]_(.4){\beta_0}|(.285)\hole
			\ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1
			\ar@{.>}@/_.35cm/[dlll]^(.4){\beta_1}
			\ar[dl]|(.28)\hole^{y_1\circ b_0}& K_1 \ar[d]^{d_1\circ
				u_1}\ar[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ
				a_1}\\G'_1 && \ar[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2 &&
			\ar[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_3}\]

>>>>

\begin{abstract}
%  When two sequences of steps executed by a computational device can
%  be considered equivalent?  This is a relevant question if one
%  considers concurrent and distributed systems, since the independence
%  of two steps means that they could be executed in any order, and
%  potentially in parallel.
  When can two sequential steps performed by a computing device be
  considered (causally) independent?  This is a relevant question for
  concurrent and distributed systems, since independence means that
  those steps could be executed in any order, and potentially in parallel.
  % 
  We investigate the issue in the context of adhesive categories, a
  very general instance of the DPO approach to rewriting. More
  precisely, we focus on left-linear rules, which not only consume,
  preserve and generate entities, as linear rules do, but may
  ``merge'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box, but it requires the
  restriction to well-behaved classes of left-linear rules.
  %
  Moreover, a standard characterisation of independence that -- in the
  linear case -- allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds.
  %and this requires a substantial revision of the theory.
  %
  The paper performs an in-depth study of the notion of independence
  for left-linear rules: it introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent computations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}

@@

So far, most of the theoretical results focussed on what are called
\emph{linear rules}: intuitively, a step is required only to consume,
preserve and generate entities.  However, in some situations it is
also necessary to ``merge'' parts of the state.  This happens already
in the context of string diagrams mentioned above, but also in
graphical implementations of nominal calculi where, as a result of
name passing, the received name is identified with a local one at the
receiver~\cite{CVY:ESSPE,Gad07} or in the visual modelling of bonding
in biological/chemical processes~\cite{PUY:MBPE}. A similar mechanisms
is at the core of e-graphs (equality graphs), recently used for
rewrite-driven compiler optimisations~\cite{WNW:egg}: rather than
explicitly merging nodes corresponding to equal expression, the idea
here -- conceptually and, as we will see, also mathematically similar --
is to maintain an equivalence %\todo{P: more precisely a congruence}
over the nodes.  
%\todo{P: la sequenza di lavori su questo \`e ampia,
%vedere se c'\`e qualcosa d'altro}

Technically, to acquire the expressiveness needed for capturing
fusions, i.e. the merge of some elements in the state, requires to
move from linear to left-linear rewriting rules. While left-linear
rules have been considered in various instances (mainly in categories
of graphs) and some results concerning the corresponding rewriting
theory have been put forward, to the best of our knowledge the
phenomena arising when devising a theory of independence and
concurrency for rewriting with left-linear rules in the general
setting of adhesive and quasi-adhesive categories
%(which capture most ``applications'') 
have not been systematically
explored~\cite{Ehrig1976,EHP:BRfTToHLRS,baldan2011adhesivity},
even if an investigation in the setting of quasi-topoi, considering general
non-linear rules yet restricting their applicability,
% of rules and focussing as far as DPO in concerned on quasi-adhesivity, 
has been presented in~\cite{BehrHK23}.

% \todo{P: see also \cite{HMP:DPORADD,BHK:CTNL} ... verificare e aggiungere}

Here we perform this exploration at the level of $\mathcal{M}$-adhesive categories, 
%i.e.such that adhesivity holds only for a class $\mathcal{M}$ of arrows, 
which encompass a large family of DPO-based formalisms adopted by researchers and 
pratictioners alike~\cite{xxx}.

@@

% \begin{enumerate}
% \item \emph{confluence of switching}: when transforming derivation
%   sequences into equivalent ones by switching independent steps, the
%   result not depends on the order in which switchings are applied, but
%   only on the resulting permutation;

% \item \emph{globality of independence}: if two independent steps are
%   moved forward or backward due to the application of switchings to a
%   derivation, they remain independent.
% \end{enumerate}

There is, however, a further problem. In the literature on linear
rewriting systems, a central result is the a characterisation of
switch equivalence of sequences of steps based on what are called
consistent permutations or processes~\cite{Handbook, tobias}.
In turn this is the key to derive some simple but fundamental properties 
for linear rewriting systems, namely
%
\begin{enumerate}
\item \emph{consistency of switching}: when transforming
  sequences of rewriting steps into equivalent ones by switching independent steps, the
  result not depends on the order in which switchings are applied, but
  only on the resulting permutation;
\item \emph{globality of independence}: if two independent steps are
  moved forward or backward due to the application of switchings to a
  sequence, they remain independent.
\end{enumerate}
Properties (1) and (2) allow one to deduce a canonical form for equivalence proofs 
between sequences of rewriting steps seen as concurrent computations.

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails: The 
existence of a consistent permutation is necessary but not
sufficient, hence it cannot be used for proving (1) and (2).
%
This fact, noted e.g. in~\cite{BaldanGM06}, leads to
asking whether these properties hold.
We show that much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weak form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality~\cite{baldan2017domains}. Finally, 
we prove that a canonical form
for switching sequences can also be recovered.

@@

\begin{definition}[Well switching rewriting systems]
A left-linear DPO rewriting system $(\X, \R)$ is \emph{well switching} if it is strong enforcing and, for every derivation $\der{D}:=\{\dder{D}_{i}\}_{i=0}^1$, there is at most one independence pair between $\dder{D}_0$ and $\dder{D}_1$.
\end{definition}

@@

\begin{restatable}[Globality of independence]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ strong enforcing DPO rewriting system.
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is an independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item
    \label{lem:indep-global-left:1}
    $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
    switchable.

  \item
    \label{lem:indep-global-left:2}
    If, moreover, $(\X, \R)$ is well switching and $S_{i_0, i_1}(\dder{D}_0)\seqind \dder{D}_2$ with an independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)\seqind S_{\alpha_0, \alpha_1}(\dder{D}_2)$.
  \end{enumerate}
\end{restatable}

@@

	\begin{figure}[h]
		\centering
	$\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){i_0}|(.7)\hole
		\ar[dr]^{h_0}&& L_1\ar@/^.8cm/[ddll]^(.2){i_1}
		\ar[dl]_{m_1}\\&K_0\ar@{>->}[dr]^{k_0}\ar[dl]_{l_0}
		\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){v_0}\ar[ur]^{r_0}&& G_1 &&
		K_1
		\ar@/^.6cm/[ddll]|(.36)\hole_(.65){v_1}\ar[dl]_{k_1}\ar@{>->}[lu]_{l_1}
		\ar[dr]^{r_1}\\L_0
		\ar@/_.6cm/[ddrr]_(.2){j_0}|(.31)\hole|(.81)\hole
		\ar[dr]^(.4){m_0}|(.61)\hole && D_0
		\ar@{>->}[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&D_1
		\ar[dr]|(.4)\hole^(.65){g_1}
		\ar@{>->}[ul]|(.5)\hole_(.65){f_1}&&R_1\ar@/^.6cm/[ddll]^(.2){j_1}|(.31)\hole|(.81)\hole\ar[dl]_(.4){h_1}|(.6)\hole\\&G_0
		&&P_1\ar[dr]_{q_1}
		\ar@{>->}[dl]^{q_0}\ar[ur]^(.4){p_1}\ar@{>->}[ul]_(.4){p_0}&&G_2\\L_1
		\ar@/^.6cm/[uurr]^(.2){i_1} \ar[ur]_(.35){f_0\circ
			i_1}|(.61)\hole&&Q_0
		\ar@{>->}[ul]|(.4)\hole_(.65){s_0}\ar[dr]|(.5)\hole^(.65){t_0}
		&&Q_1\ar[ur]|(.4)\hole^(.65){t_1} \ar@{>->}[dl]|(.5)\hole_(.65){s_1}
		&& R_0 \ar[ul]^(.35){g_1\circ
			i_0}|(.61)\hole\ar@/_.6cm/[uull]_(.2){i_0}\\&K_1
		\ar[ur]_{q_0\circ v_1} \ar[dr]_{r_1}
		\ar@{>->}[ul]^{l_1}\ar@/^.6cm/[uurr]_(.65){v_1}&&H'_1&& K_0
		\ar[ur]_{r_0} \ar[ul]^{q_1\circ v_0} \ar@{>->}[dl]^{l_0}
		\ar@/_.6cm/[uull]^(.65){v_0}\\&& R_1
		\ar[ur]_{\hspace{-6pt}s_1\circ
			j_1}\ar@/^.8cm/[uurr]^(.2){j_1}&& L_0 \ar[ul]^{t_0\circ
			j_0\hspace{-6pt}} \ar@/_.8cm/[uull]_(.2){j_0} |(.69)\hole
	}	$
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switch of $\dder{D}_0$ and $\dder{D}_1$}
		\label{fi:first}
	\end{figure}

	\begin{figure}[h]
		\centering 
		$\xymatrix@C=15pt@R=15pt{&&R_1 \ar@/_.8cm/[ddrr]_(.2){a_0}|(.7)\hole
			\ar[dr]^{h_1}&& L_2\ar@/^.8cm/[ddll]^(.2){a_1}
			\ar[dl]_{m_2}\\&K_1\ar[dr]^{k_1}\ar@{>->}[dl]_{l_1}
			\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){u_1}\ar[ur]^{r_1}&& G_2 &&
			K_2
			\ar@/^.6cm/[ddll]|(.36)\hole_(.65){u_2}\ar[dl]_{k_2}\ar@{>->}[lu]_{l_2}
			\ar[dr]^{r_2}\\L_1
			\ar@/_.6cm/[ddrr]_(.2){b_0}|(.31)\hole|(.81)\hole
			\ar[dr]^(.4){m_1}|(.61)\hole && D_1
			\ar@{>->}[dl]|(.4)\hole_(.65){f_1}\ar[ur]|(.5)\hole^(.7){g_1}&&D_2
			\ar[dr]|(.4)\hole^(.65){g_2}
			\ar@{>->}[ul]|(.5)\hole_(.65){f_2}&&R_2\ar@/^.6cm/[ddll]^(.2){b_1}|(.31)\hole|(.81)\hole\ar[dl]_(.4){h_2}|(.61)\hole\\&G_1
			&&P_2\ar[dr]_{d_1}
			\ar@{>->}[dl]^{d_0}\ar[ur]^(.4){c_1}\ar@{>->}[ul]_(.4){c_0}&&G_3\\L_2
			\ar@/^.6cm/[uurr]^(.2){a_1} \ar[ur]_(.35){f_1\circ
				a_1}|(.61)\hole&&Q_2
			\ar@{>->}[ul]|(.4)\hole_(.65){x_1}\ar[dr]|(.5)\hole^(.65){y_1}
			&&Q_3\ar[ur]|(.4)\hole^(.65){y_2} \ar@{>->}[dl]|(.5)\hole_(.65){x_2}
			&& R_1 \ar[ul]^(.3){g_2\circ
				a_0}|(.61)\hole\ar@/_.6cm/[uull]_(.2){a_0}\\&K_2
			\ar[ur]_{d_0\circ u_2} \ar[dr]_{r_2}
			\ar@{>->}[ul]^{l_2}\ar@/^.6cm/[uurr]_(.65){u_2}&&G'_2&& K_1
			\ar[ur]_{r_1} \ar[ul]^{d_1\circ u_1} \ar@{>->}[dl]^{l_1}
			\ar@/_.6cm/[uull]^(.65){u_1}\\&& R_2
			\ar[ur]_{\hspace{-6pt}x_2\circ
				b_1}\ar@/^.8cm/[uurr]^(.2){b_1}&& L_1 \ar[ul]^{y_1\circ
				b_0\hspace{-7pt}} \ar@/_.8cm/[uull]_(.2){b_0} |(.69)\hole
		}	$		
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switch of $\dder{D}_1$ and $\dder{D}_2$}
		\label{fi:second}
	\end{figure}

	\begin{figure}
		\centering
		$\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){e_0}|(.7)\hole
			\ar[dr]^{\hspace{-5pt}h_0}&& L_2\ar@/^.8cm/[ddll]^(.2){e_1}
			\ar[dl]_{f_1\circ a_1\hspace{-5pt}}\\&K_0\ar[dr]^{k_0}\ar@{>->}[dl]_{l_0}
			\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){w_0}\ar[ur]^{r_0}&& G_1 &&
			K_2 \ar@/^.6cm/[ddll]|(.36)\hole_(.65){w_2}\ar[dl]_{d_0\circ
				u_2\hspace{-5pt}}\ar@{>->}[lu]_{l_2} \ar[dr]^{r_2}\\L_0
			\ar@/_.6cm/[ddrr]_(.2){o_0}|(.31)\hole|(.81)\hole
			\ar[dr]^(.4){m_0}|(.61)\hole && D_0
			\ar@{>->}[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&Q_2
			\ar[dr]|(.4)\hole^(.65){y_1}
			\ar@{>->}[ul]|(.5)\hole_(.65){x_1}&&R_2\ar@/^.6cm/[ddll]^(.2){o_1}|(.31)\hole|(.81)\hole\ar[dl]_(.3){x_2\circ
				b_1}|(.58)\hole \\&G_0 &&P_3\ar[dr]_{n_1}
			\ar@{>->}[dl]^{n_0}\ar[ur]^(.4){\theta_1}\ar@{>->}[ul]_(.4){\theta_0}&&G'_2\\L_2
			\ar@/^.6cm/[uurr]^(.2){e_1} \ar[ur]_(.35){f_0\circ
				e_1}|(.61)\hole&&Q_4
			\ar@{>->}[ul]|(.4)\hole_(.65){z_0}\ar[dr]|(.5)\hole^(.65){z'_0}
			&&Q_5\ar[ur]|(.4)\hole^(.65){z'_1}
			\ar@{>->}[dl]|(.5)\hole_(.65){z_1} && R_0 \ar[ul]^(.3){y_1\circ
				e_0}|(.6)\hole\ar@/_.6cm/[uull]_(.2){e_0}\\&K_2
			\ar[ur]_{\hspace{-2pt}n_0\circ w_2} \ar[dr]_{r_2}
			\ar@{>->}[ul]^{l_2}\ar@/^.6cm/[uurr]_(.65){w_2}&&G'_1&& K_0
			\ar[ur]_{r_0} \ar[ul]^{n_1\circ w_0} \ar[dl]^{l_0}
			\ar@/_.6cm/[uull]^(.65){w_0}\\&& R_2
			\ar[ur]_{\hspace{-4pt}z_1\circ
				o_1}\ar@/^.8cm/[uurr]^(.2){o_1}&& L_0 \ar[ul]^{z'_0\circ
				o_0\hspace{-7pt}} \ar@/_.8cm/[uull]_(.2){o_0} |(.69)\hole
		}	$
		\caption{Picture for the proof of \Cref{lem:indep-global-left}: switch of $\dder{D}_0$ and $S_{a_0,a_1}(\dder{D}_2)$}
		\label{fi:third}
	\end{figure}

	\begin{enumerate}
		\item We have to construct the two dotted arrows in the diagram
		below.
		\[\xymatrix@C=15pt{L_0 \ar[d]_{z'_0\circ o_0}&& K_0
			\ar[d]_{n_1\circ w_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
			\ar@{.>}@/^.35cm/[drrr]_(.4){\beta_0}|(.285)\hole
			\ar[dr]|(.28)\hole_{y_1\circ e_0} && L_1
			\ar@{.>}@/_.35cm/[dlll]^(.4){\beta_1}
			\ar[dl]|(.28)\hole^{y_1\circ b_0}& K_1 \ar[d]^{d_1\circ
				u_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{g_2\circ
				a_1}\\G'_1 && \ar@{>->}[ll]^{z_1} Q_5 \ar[rr]_{z'_1}&& G'_2 &&
			\ar@{>->}[ll]^{x_2} Q_3 \ar[rr]_{y_2}&& G_3}\]
